A Streaming Algorithm for $2$-Center with Outliers in High Dimensions

### 2015-04-26 paper - Copy.tex
paper reorganized
Passed to Behnam for completing the proofs, etc.

\usepackage{graphicx}
\usepackage{caption,subcaption}

\usepackage[noend]{algpseudocode}

%----------------------- Environments --------------------------
\newtheorem{corollary}{Corollary} 
\newtheorem{problem}{Problem}
\newtheorem{algorithm}{Algorithm}
\newtheorem{definition}{Definition}

%------------------------ Paper Notations -----------------------------
\DeclareMathOperator*{\argmin}{arg\,min}

%-------------------------- Notations ------------------------------------
\newcommand{\cO}{\ensuremath{{\rm O}}}
\newcommand{\st}{\textsuperscript{\textit{st}} }
\newcommand{\nd}{\textsuperscript{\textit{nd}} }
\newcommand{\nth}{\textsuperscript{\textit{th}} }

@@

\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\seq}[1]{\left< #1 \right>}
\newcommand{\ceil}[1]{\left\lceil{#1}\right\rceil}
\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
\newcommand{\card}[1]{\left|{#1}\right|}
\newcommand{\setcomp}[1]{\overline{#1}}
\newcommand{\provided}{\,:\,}

@@

%----------------------- Title -----------------------------------------
\title{Streaming $2$-Center with Outliers in High Dimension}
\author{Hamid~Zarrabi-Zadeh\thanks{Department of Computer Engineering, 
	Sharif University of Technology.
	{\tt zarrabi@sharif.edu}}
	\and 
}

%------------------------------ Text -------------------------------------
\begin{document}

@@

\begin{abstract}
In this paper, we study the $2$-center streaming problem with $z$ outliers in a Euclidean space, which means finding two congruent balls covering all the points except $z$ outliers in a streaming work-space. Before considering this problem, we provide a  $2$-approximation algorithm for $1$-center with $z$ outliers in a general space. Our algorithm for the $2$-center problem uses an update time and working space polynomial in $d$ and sub-linear in $n$ to calculate a $(1.8+\eps)$-approximation solution.
\end{abstract}

@@

\section{Introduction}
Metric $k$-center problem -- covering a set of point with $k$ congruent balls of minimum radius -- is a fundamental problem in computer science and has various applications in data mining and image processing.
One of $k$\nobreakdash-center problem variants is $k$-center with $z$-outlier which finds $k$ congruent balls with minimum radius covering all but $z$ points of a point set. In real world applications, due to commonness of noisy data, it is important to consider outlier points. Because even a small number of outliers ($\cO(k)$) can increase the radius of $k$\nobreakdash-center problem unboundedly.
In this paper we focus on streaming data model. In the streaming data model, input data arrives as a sequence and there is not enough space to keep the whole data set. Algorithms in this model have limitations in space (typically sub-linear in input size) and processing time per item. This model is specially suitable for massive data sets, due to its sub-linear space complexity \cite{aggarwal2007data}.

@@

\section{Preliminaries}
\label{sec:pre}
Let $B(c,r)$ denote a ball with center $c$ and radius $r$. We denote by $pq$ the straight line segment between these two points and by $\card{pq}$ the length of this segment. 
Define $\delta$ of two balls $B(c,r)$ and $B'(c',r')$, the minimum distance between $B$ and $B'$ which is equal to $\max (0, \card{cc'}-r-r')$.

Let $B_{2,1}^*(c_1^*, r_2^*)$ and $B_{2,2}^*(c_2^*, r_2^*)$ denote the optimal solution to $2$-center problem with $z$ outliers. By $\delta^*$, denote the distance between $B_{2,1}^*$ and $B_{2,2}^*$, that is, $\delta^* = \max (0, \card{c_1^*c_2^*} - 2r^*)$.

Two congruent balls $B_1$ and $B_2$ with radius $r$ are said to be \emph{$C$-separated}, whenever the distance $\delta$ of two balls is at least $Cr$.

%Let $\mathcal{P}$ be a configuration of $n$ point in $d$-dimensional Euclidean space. A point $c_p$, not necessarily in $\mathcal{P}$, is a \emph{centerpoint} of $\mathcal{P}$, if no open half-space that omits $c_p$ contains more than $\frac{dn}{d+1}$ points of $\mathcal{P}$.

Given an $n$-point set $\mathcal{P}$ in $d$-dimensional Euclidean space, a point $c \in \IR^d$ is called a \emph{centerpoint} of $\mathcal{P}$ if any halfspace containing $c$, contains at least $\ceil{\frac{n}{d + 1}}$ points of $\mathcal{P}$. In other words, any halfspace (or convex set) that avoids a centerpoint can contain at most $\floor{\frac{dn}{d + 1}}$ points of $\mathcal{P}$.

@@

\section{A Simple $2$-Approximation Algorithm for $1$-Center with Outliers}
\label{sec:2apr}

\begin{algorithm}
\label{alg:1center}
\leavevmode
\begin{algorithmic}
\Function{$1$-centerWithOutliers}{$\mathcal{P}, z$}
	\State Let $p_i$ denote the elements of $\mathcal{P}$, for $i \in \set{1, \dots, n}$.
	\ForAll{$i \gets 1, \dots, z+1$}
		\State Assume $p_i$ is a non-outlier point in optimal solution.
		\State $B(c_i, r_i) \gets$ \Call{KnownPoint}{$\mathcal{P}, z, p_i$}
	\EndFor
	\State $i^* \gets \argmin_{i=1}^{z+1} r_i$
	\State \Return $B(c_{i^*}, r_{i^*})$
\EndFunction
\Statex
\Function{KnownPoint}{$\mathcal{P}, z, p$}
	\State $B \gets B(p, 0)$
	\State $Q \gets \emptyset$
	\For{$p'$ in $\mathcal{P}$}
		\If{$p' \notin B$}
			\State Insert $p'$ into $Q$.
			\If{$\card{Q} = z + 1$}
				\State $q$ $\gets$ closest point to $p$ in $Q$
				\State Remove $q$ from $Q$.
				\State $B \gets B(p, \card{pq})$
			\EndIf
		\EndIf
	\EndFor
	\State \Return{B}
\EndFunction
\end{algorithmic}
\end{algorithm}
Function \textproc{KnownPoint} in Algorithm~\ref{alg:1center}, takes a point $p$ as an argument which is guaranteed to be a non-outlier point in the optimal solution. To overcome the lack of knowledge about such a point, function \textproc{$1$\nobreakdash-centerWithOutliers} tries each of the first $z+1$ points of the stream as a candidate for a non-outlier point and generates a ball for each, using function \textproc{KnownPoint}. It returns the ball with the minimum radius. Note that, due to space limitations inherent to streaming data model, the body of for-all loop in function \textproc{$1$-centerWithOutliers} runs in parallel.

\begin{theorem}
Algorithm~\ref{alg:1center} is a $2$-approximation algorithm for $1$-center problem with $z$ outliers, in any dimension.
\end{theorem}

\begin{proof}
Let $B^*(c^*, r^*)$ be the optimal solution. Clearly, there exists a point $p$ among the first $z+1$ points of $\mathcal{P}$, which is not an outlier in the optimal solution (Figure~\ref{fig:1center}). Since $p \in B^*$, then $\card{pp'} \le 2r^*$ for any point $p'\in B^*$. There is at least one point $q'$ among the $(z+1)$-furthest points from $p$ that is not an outlier in the optimal solution. Thus $\card{pq} \le \card{pq'} \le 2r^*$ and Algorithm~\ref{alg:1center} returns a valid solution of radius at most $2r^*$. So the proof is complete.
\end{proof}

%
%\begin{figure}[h]
%	\centering
%	\includegraphics[width=17em]{figs/one-center}
%	\caption{Output of Algorithm~\ref{alg:1center}}
%	\label{fig:1center}
%\end{figure}

% Behnam: Aks be nazaram gooya nist ziad. va inke jaii behesh refrence dade nashode!
% Kiana: hagh ba behname. akse bedune refrence nadarim asan. montaha man benazaram kheili ehtiaji nis behesh asan 
% Sahand: hazf konim ya pishnehadi baray e behboodesh darid?
% Kiana: Man be nazaram vaghean chon kheili az idehaye basic ham baraye gheire streaming hamine va osulan idash yekam badihitowre hazf beshe. felan commentesh kardam age nazaretun chize digei bud avazesh konin.

The space complexity of Algorithm~\ref{alg:1center} is $\cO(z^2 + zd)$, the update time is $\cO (z \log z + zd)$ and the query time is $\cO (z)$.
If a non-outlier point $p$ is known then the space, update time and query time complexities of Algorithm~\ref{alg:1center} can be reduced by a factor of $z$.

@@

\section{The $2$-Center Problem with $z$ Outliers}

In all the algorithms given in this section it is assumed that $p_1$ is a non-outlier point. It is not hard to see that, this limitation can be overcome by considering $\cO(z)$ parallel instances of the algorithm, similar to Algorithm~\ref{alg:1center}.
% Behnam: Move this to upper section, this assumption can be used for previous algorithm
%kiana: Here is better. Since we do not have p_1 in section 4.2

@@

In Section \ref{subsec:smaller}, we give a description of our algorithm for a given $r' > 0$ for the case where $\delta^* < C r^*$ and $1.2r^* \le r' < (1.2 + 2\eps/3)r^*$, which returns a $(1.8 + \eps)$-approximate solution. We explain how to find such an $r'$ and present a full description of our algorithm for the case where $\delta^* < C r^*$ in Subsection~\ref{subsec:findr}. Then we inspect the case where $\delta^* \ge C r^*$ in Subsection~\ref{subsec:bigger}.

\subsection{The Case $\delta^* < C r*$}
\label{subsec:smaller}
Our idea in this section is derived from \cite{ahn2014computing, kim2014improved}. To avoid duplication, we just outline the important parts of their algorithm and our modifications to it. Kim's algorithm has $10$ different states. In each step one or many of the states can be valid and their algorithm considers all of them in parallel. In each state, there exist at most two balls. A transition between the states occurs whenever a point not covered by any of the two balls has arrived in the stream. This transition graph starts with a blank node (with no balls) and has depth four (without considering the end nodes). It is straightforward to see that this transition graph is a DAG.
Our modification is on the transition part.
Points that are covered by current solution can safely be ignored, as they do not cause any transitions in the state, and only those points that lie outside the solution are candidates for being an outlier.
In each state the number of points that should be considered as outliers is unknown. So all the possible choices should be considered. It is sufficient to have four integers, $n_1,\dots, n_4$ representing the number of outliers in depth $1$ to $4$, such that $\sum_{i=1}^{4} n_i=z$. To sum up, the modified algorithm is defined as follows:
{
\algrenewcommand\algorithmicindent{0.8em}%
\begin{algorithm}
\label{alg:smallc}
\leavevmode
\begin{algorithmic}
\Function{Known$r'$}{$\mathcal{P}, z, r'$}
	\For{$(n_1, \dots, n_4) \gets \set{(n_1, \dots, n_4) \provided \sum n_i=z}$}
		\State $CS \gets B(p_1,r')$
		\State $p_{o_2} \gets $ \Call{NthOutlier}{$n_1 + 1$, $p_1$}
		\If{$p_{o_2}$ arrives}
			\State $CS\gets [ \text{the sub-case of $p_{o_2} \in B_{2, 1}^*$} ] \cup [ \text{the sub-case of $p_{o_2} \in B_{2,2}^*$} ]$
			\For{each case $S$}
				\State $p_{o_3} \gets$ \Call{NthOutlier}{$n_2 + 1$, $p_{o_2}$}
				\If{$p_{o_3}$ arrives} 
					\State $CS\gets CS\setminus S \cup [ \text{the sub-case of $p_{o_3} \in B_{2, 1}^*$} ] \cup [ \text{the sub-case of $p_{o_3} \in B_{2,2}^*$}]$
					\For {each sub-case $S'$} 
						\State $p_{o_4} \gets$ \Call{NthOutlier}{$n_3 + 1$, $p_{o_3}$}
						\If {$p_{o_4}$ arrives} 
							\State Replace $S'$ with a new candidate solution.
							\State $p_{o_5} \gets$ \Call{NthOutlier}{$n_4 + 1$, $p_{o_4}$}
							\If{$p_{o_5}$ arrives}
								\State Abandon the solution.
							\EndIf
						\EndIf
					\EndFor
				\EndIf
			\EndFor

		\EndIf
		\State $\mathop{Sol}_{n_1,n_2,n_3,n_4} \gets$ the solution with the smallest larger radius, among all the sub-cases.
	\EndFor
	\State \Return the solution $\mathop{Sol}_{n_1,n_2,n_3,n_4}$ with the smallest radius, for all valid values of $(n_1,n_2,n_3,n_4)$.
\EndFunction
\end{algorithmic}
\end{algorithm}
}
Here, \textproc{NthOutlier}$(n,p)$ returns the $n$\nth point arriving after $p$ in $\mathcal{P}$ that does not lie in current candidate solution.

\begin{theorem}
Given $\delta^* < C r^*$ and $1.2r^* \le r' < (1.2 + 2\eps/3)r^*$, Algorithm~\ref{alg:smallc} uses $\cO (dz^3)$ space and returns a $(1.8+\eps)$-approximate solution. Algorithm~\ref{alg:smallc} spends $\cO (dz^3)$ update time for each point in $\mathcal{P}$ and answers a query in $\cO (dz^3)$.
\end{theorem}

\begin{proof}
Since our algorithm considers all possible input cases of streaming points, there is at least one feasible solution. Since every feasible solution has its larger radius at most $3r'/2$, the final solution has larger radius at most $3r'/2 \le (1.8 + \eps)r^*$.
For space complexity, our algorithm maintains at most two balls in each case, and therefore it uses $\cO(dz^3)$ space. Whenever the next point is inserted, the algorithm updates the solution for each sub-case in $\cO (d)$ time. Therefore, the algorithm spends $\cO (dz^3)$ update time for each point of $\mathcal{P}$. Answering a query consists of choosing the minimum radius among all the candidate solutions, which amounts to $\cO (dz^3)$ time.
\end{proof}

@@

\begin{lemma}
\label{lem:2lt1}
Let $r_1^*, r_2^*$ be the radius of $1$-center and $2$-center optimal solutions with $z$ outliers, respectively. Then $r_1^* < \left(\frac{C}{2} + 2\right) r_2^*$.
\end{lemma}
\begin{proof}
Suppose that $(B_{2, 1}^*(c_1^*, r_2^*), B_{2,2}^*(c_2^*, r_2^*))$ is an optimal solution for $2$-center problem. Consider $m$ as the midpoint of segment 
$c_1^*c_2^*$ (Figure~\ref{fig:2lt1}).
Clearly, $B\left(m, \frac{\delta}{2} + 2r_2^*\right)$ covers both $B_{2, 1}^*$ and $B_{2,2}^*$. So it is a solution for $1$-center problem and the proof is complete.
\end{proof}

@@

Algorithm~\ref{alg:1center} finds a $2$-approximation for $r_1^*$. By Lemma \ref{lem:1lt2} and \ref{lem:2lt1}, it gives a $2\left(\frac{C}{2} + 2\right)$-approximation for $r_2^*$.

@@

Our algorithm maintains $m=\ceil{\frac{1.2(3C+12)}{\eps}}$ candidate lengths.
For each candidate length $t$, assume that $r' = t$ and use it to run Algorithm~\ref{alg:smallc}. Algorithm~\ref{alg:1center} is used to estimate $r'$. Let $k_i$ be the $i$\nth non-zero radius calculated by the algorithm. Obviously, the sequence of answers given by the algorithm is increasing.

Let $l_1=k_1$ and $l_i=2^j l_1$, where $l_i$ is the smallest number satisfying $l_i \ge k_i$. By Observation~\ref{obs:2apr}, $l_i$ is a $(2C + 8)$-approximation for $r_2^*$. If the interval $(0, 1.2 l_i]$ divides into $m$ equal segments (Figure~\ref{fig:findr}), then length of each segment is at most $2/3 \eps r_2^*$ and the endpoints as candidates are $\mathcal{L}_i = \set{j \times t_i \provided j = 1, \dots, m}$ where $t_i = \frac{1.2 l_i}{m}$. 
The inequality $1.2 r_2^* \le j \times l_i \le (1.2 + \frac{2\eps}{3})r_2^*$ holds for at least one of these candidates.

@@

\subsection{The Case $\delta^* \ge C r^*$}
\label{subsec:bigger}

\begin{obs}
\label{obs:c+4}
Let $d$ be the distance between arbitrary pair of points from two $C$-separated balls. Then $1 \le \frac{d}{\delta} \le \frac{C+4}{C}$.
\end{obs}

@@

\begin{cor}
\label{cor:omitting-centerpoint}
Given a set $\mathcal{P}$ of $k(d+1)$ points in $d$\nobreakdash-dimensional Euclidean space, there is a point $c_p$, not necessarily in $\mathcal{P}$, that any convex object that does not cover $c_p$, leaves at least $k$ points of $\mathcal{P}$ uncovered.
\end{cor}

>>>>

\usepackage{caption,subcaption,cleveref}

\usepackage{algorithm}
\usepackage[noend]{algorithmic}

%----------------------- Environments --------------------------

\newtheorem{corollary}{Corollary} 
\newtheorem{problem}{Problem}
%\newtheorem{algorithm}{Algorithm}
\newtheorem{definition}{Definition}

%------------------------ Algorithm -----------------------------

\newcommand{\textproc}{\textsc}
\newcommand{\Call}[2]{\textsc{#1}(#2)}
\renewcommand{\algorithmicforall}{{\bf for each}}

%------------------------ Paper Notations -----------------------------

\DeclareMathOperator*{\argmin}{arg\,min}

%-------------------------- Notations ------------------------------------

\newcommand{\cO}{\ensuremath{{\rm O}}}
\newcommand{\st}{\textsuperscript{\textit{st}} }
\newcommand{\nd}{\textsuperscript{\textit{nd}} }
\newcommand{\nth}{\textsuperscript{\textit{th}} }

@@

\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\seq}[1]{\left< #1 \right>}
\newcommand{\ceil}[1]{\left\lceil{#1}\right\rceil}
\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
\newcommand{\card}[1]{\left|{#1}\right|}
\newcommand{\len}[1]{\|{#1}\|}
\newcommand{\setcomp}[1]{\overline{#1}}
\newcommand{\provided}{\,:\,}

@@

%----------------------- Title -----------------------------------------

\title{A Streaming Algorithm for $2$-Center with Outliers in High Dimensions}
\author{Behnam~Hatami\thanks{Department of Computer Engineering, 
	Sharif University of Technology.
	{\tt hatami@ce.sharif.edu}}
	\and 
	Hamid~Zarrabi-Zadeh\thanks{Department of Computer Engineering, 
	Sharif University of Technology.
	{\tt zarrabi@sharif.edu}}
}

%------------------------------ Text -------------------------------------

\begin{document}

@@

\begin{abstract}
We study the Euclidean $2$-center problem with outliers in the data stream model. 
Given a stream of points in arbitrary $d$ dimensions, the goal is to find two congruent balls 
of minimum radius covering all but $z$ points. 
We provide a $(1.8+\eps)$-approximation streaming algorithm for the problem, %2-center with outliers,
improving upon the previous $(3 + \eps)$-approximation algorithm available for the problem.
The space complexity and update time of our algorithm is $\poly(d, z, 1/\eps)$,
independent of the size of the stream.
\end{abstract}

@@

\section{Introduction}
The $k$-center problem---covering a set of points 
using $k$ congruent balls of minimum radius---is a fundamental problem 
in computer science, having many applications in various areas
such as data mining, unsupervised learning, statistics, and image processing.
One of $k$\nobreakdash-center problem variants is $k$-center with $z$-outlier which finds $k$ congruent balls with minimum radius covering all but $z$ points of a point set. In real world applications, due to commonness of noisy data, it is important to consider outlier points. Because even a small number of outliers ($\cO(k)$) can increase the radius of $k$\nobreakdash-center problem unboundedly.
In this paper we focus on streaming data model. In the streaming data model, input data arrives as a sequence and there is not enough space to keep the whole data set. Algorithms in this model have limitations in space (typically sub-linear in input size) and processing time per item. This model is specially suitable for massive data sets, due to its sub-linear space complexity \cite{aggarwal2007data}.

@@

\section{Preliminaries}
\label{sec:pre}
Let $B(c,r)$ denote a ball of radius $r$ centered at $c$. 
We denote by $pq$ the straight line segment between $p$ and $q$,
and denote by $\len{pq}$ the length of this segment. 
Given two balls $B(c,r)$ and $B'(c',r')$, 
we define their \emph{distance} to be $\delta = \max \set{0, \len{cc'}-r-r'}$.

Let $B_{2,1}^*(c_1^*, r_2^*)$ and $B_{2,2}^*(c_2^*, r_2^*)$ denote the optimal solution to $2$-center problem with $z$ outliers. By $\delta^*$, denote the distance between $B_{2,1}^*$ and $B_{2,2}^*$, that is, $\delta^* = \max (0, \len{c_1^*c_2^*} - 2r^*)$.

Two congruent balls $B_1$ and $B_2$ with radius $r$ are said to be \emph{$\alpha$-separated}, whenever the distance $\delta$ of two balls is at least $Cr$.

%Let $P$ be a configuration of $n$ point in $d$-dimensional Euclidean space. A point $c_p$, not necessarily in $P$, is a \emph{centerpoint} of $P$, if no open half-space that omits $c_p$ contains more than $\frac{dn}{d+1}$ points of $P$.

Given an $n$-point set $P$ in $d$-dimensional Euclidean space, a point $c \in \IR^d$ is called a \emph{centerpoint} of $P$ if any halfspace containing $c$, contains at least $\ceil{\frac{n}{d + 1}}$ points of $P$. In other words, any halfspace (or convex set) that avoids a centerpoint can contain at most $\floor{\frac{dn}{d + 1}}$ points of $P$.

@@

\section{A $2$-Approximation for $1$-Center with Outliers}
\label{sec:1-center}

We warm up by presenting a simple $2$-approximation streaming algorithm
for the $1$-center problem with outliers.
The pseudocode is provided in \Cref{alg:1-center}.
It receives as input a stream of points, $P$,
and a special point $c \in P$ which is guaranteed to be 
a non-outlier in the optimal solution.
The algorithm returns a ball $B$ containing all but $z$ points of $P$.
%We will show later how to remove this assumption.

%\begin{algorithm}
%\caption{\sc 1-Center$(P, z)$} 
%\algsetup{indent=1.5em}
%\label{alg:1-center}
%\begin{algorithmic}[1]
%
%	%\STATE Let $p_i$ denote the elements of $P$, for $i \in \set{1, \dots, n}$.
%	\STATE let $p_1, \ldots, p_n$ be the points of $P$
%	\FOR{$i = 1, \dots, z+1$}
%		%\STATE move $p_i$ to the front of $P$ %is a non-outlier point in optimal solution.
%		\STATE $B_i \gets$ \Call{1-Center-Special}{$P, z,p_i$}
%	\ENDFOR
%	%\STATE $i^* \gets \argmin_{i=1}^{z+1} r_i$
%	\RETURN smallest ball among $\set{B_1, \ldots, B_{z+1}}$
%
%\end{algorithmic}
%\end{algorithm}

\begin{algorithm}
\caption{\sc 1-Center$(P, c, z)$} 
\algsetup{indent=1.5em}
\label{alg:1-center}
\begin{algorithmic}[1]
	%\STATE {$\triangleright$ $c$ is a non-outlier point in the optimal solution}
	\STATE $B \gets B(c, 0)$
	\STATE $Q \gets \emptyset$
	\FORALL {$p$ in $P$}
		\IF {$p \notin B$}
			\STATE insert $p$ into $Q$
			\IF {$|Q| = z + 1$}
				\STATE $q$ $\gets$ closest point to $c$ in $Q$
				\STATE remove $q$ from $Q$
				\STATE $B \gets B(c, \len{cq})$
			\ENDIF
		\ENDIF
	\ENDFOR
	\RETURN {B}
\end{algorithmic}
\end{algorithm}

%The function \textproc{1-Center-Special} in \Cref{alg:1-center-special}, 
%takes as input a point $c$ which is guaranteed to be a non-outlier point in the optimal solution. 
%To overcome the lack of knowledge about such a point, 
%function \textproc{1-Center} in \Cref{alg:1-center}
%tries each of the first $z+1$ points of the stream as a candidate for a non-outlier point and 
%generates a ball for each, using function \textproc{1-Center-Special}. 
%It returns the ball with the minimum radius. 
%Note that, in the streaming model, %due to space limitations inherent to streaming data model, 
%the body of the for loop in function \textproc{1-Center} runs in parallel.
%To run function \textproc{1-Center} in the data stream model,
%we run the body of the for loop in $z+1$ parallel streaming instances.

\begin{theorem} \label{thm:1-center-special}
	Algorithm~\ref{alg:1-center} computes a $2$-approximation to
	the $1$-center problem with $z$ outliers,
	provided that a non-outlier point in the optimal solution in given.
\end{theorem}

\begin{proof}
Let $B^*(c^*, r^*)$ be the optimal solution,
and $c$ be a non-outlier point in the optimal solution.
Since $c$ is covered by $B^*$,
for all points $p \in B^*$, we have
$\len{cp} \le \len{cc^*} + \len{c^*p} \le 2r^*$.
Among the $z+1$ points furthest from $c$,
there is at least one point $q$ which is not outlier in the optimal solution,
and therefore, is contained in $B^*$
(see \Cref{fig:1center}). 
Thus, $\len{cq} \le 2r^*$,
and therefore, the ball $B(c, \len{cq})$ returned by Algorithm~\ref{alg:1-center} 
is a 2-approximation. 
\end{proof}

\begin{figure}[t]
	\centering
	\includegraphics[width=12em]{figs/one-center}
	\caption{Proof of \Cref{thm:1-center-special}}
	\label{fig:1center}
\end{figure}

In \Cref{alg:1-center}, 
we had an assumption that a point $c$ is known in advance
which is guaranteed to be a non-outlier in the optimal solution. 
To overcome the lack of knowledge about such a point, 
we run $z+1$ instances of \Cref{alg:1-center} in parallel,
each of which is given as $c$ one of the first $z + 1$ points in the stream.
Clearly, there exists a point among the first $z+1$ points of $P$, 
which is not an outlier in the optimal solution.
Therefore, the smallest ball among the $z+1$ balls computed in parallel
is always within factor 2 of the optimal solution. % by \Cref{thm:1-center-special}.
Since the space complexity of Algorithm~\ref{alg:1-center} for one instance is $\cO(zd)$, 
and its update time is $\cO (zd \log z)$, % and the query time is $\cO (z)$.
we get the following.

\begin{theorem} \label{thm:1-center}
	Given a stream of points in $d$ dimensions,
	we can maintain a 2-approximation to the 1-center with $z$ outliers
	in $\cO(z^2d)$ space and $\cO (z^2d\log z)$ update time.
\end{theorem}

%function \textproc{1-Center} in \Cref{alg:1-center}
%tries each of the first $z+1$ points of the stream as a candidate for a non-outlier point and 
%generates a ball for each, using function \textproc{1-Center-Special}. 
%It returns the ball with the minimum radius. 
%Note that, in the streaming model, %due to space limitations inherent to streaming data model, 
%the body of the for loop in function \textproc{1-Center} runs in parallel.
%To run function \textproc{1-Center} in the data stream model,
%we run the body of the for loop in $z+1$ parallel streaming instances.
%
%The space complexity of Algorithm~\ref{alg:1center} is $\cO(z^2 + zd)$, the update time is $\cO (z \log z + zd)$ and the query time is $\cO (z)$.
%If a non-outlier point $p$ is known then the space, update time and query time complexities of Algorithm~\ref{alg:1center} can be reduced by a factor of $z$.

@@

\section{The $2$-Center Problem with $z$ Outliers}
\label{sec:2-center}

In this section, we provide a $(1.8 + \eps)$-approximation algorithm
for the 2-center with outliers.
In all algorithms presented in this section,
we assume that the first point of the stream, $p_1$, is a non-outlier. 
This assumption can be easily removed by considering $z + 1$ parallel instances of the algorithm, 
similar to what we did in \Cref{sec:1-center}.
% Behnam: Move this to upper section, this assumption can be used for previous algorithm
%kiana: Here is better. Since we do not have p_1 in section 4.2

@@

In Section \ref{subsec:smaller}, we give a description of our algorithm for a given $r' > 0$ for the case where $\delta^* < \alpha r^*$ and $1.2r^* \le r' < (1.2 + 2\eps/3)r^*$, which returns a $(1.8 + \eps)$-approximate solution. We explain how to find such an $r'$ and present a full description of our algorithm for the case where $\delta^* < \alpha r^*$ in Subsection~\ref{subsec:findr}. Then we inspect the case where $\delta^* \ge \alpha r^*$ in Subsection~\ref{subsec:bigger}.

\subsection{The Case $\delta^* < \alpha r*$}
\label{subsec:smaller}
Our idea in this section is derived from \cite{ahn2014computing, kim2014improved}. To avoid duplication, we just outline the important parts of their algorithm and our modifications to it. Kim's algorithm has $10$ different states. In each step one or many of the states can be valid and their algorithm considers all of them in parallel. In each state, there exist at most two balls. A transition between the states occurs whenever a point not covered by any of the two balls has arrived in the stream. This transition graph starts with a blank node (with no balls) and has depth four (without considering the end nodes). It is straightforward to see that this transition graph is a DAG.
Our modification is on the transition part.
Points that are covered by current solution can safely be ignored, as they do not cause any transitions in the state, and only those points that lie outside the solution are candidates for being an outlier.
In each state the number of points that should be considered as outliers is unknown. So all the possible choices should be considered. It is sufficient to have four integers, $n_1,\dots, n_4$ representing the number of outliers in depth $1$ to $4$, such that $\sum_{i=1}^{4} n_i=z$. To sum up, the modified algorithm is defined as follows:

\begin{algorithm}
\caption{\sc Known$r'(P, z, r')$} 
\algsetup{indent=1em}
\label{alg:smallc}
%\leavevmode
\begin{algorithmic}[1]
%\Function{Known$r'$}{$P, z, r'$}
	\FOR{$(n_1, \dots, n_4) \gets \set{(n_1, \dots, n_4) \provided \sum n_i=z}$}
		\STATE $CS \gets B(p_1,r')$
		\STATE $p_{o_2} \gets $ \Call{NthOutlier}{$n_1 + 1$, $p_1$}
		\IF{$p_{o_2}$ arrives}
			\STATE $CS\gets [ \text{the sub-case of $p_{o_2} \in B_{2, 1}^*$} ] \cup [ \text{the sub-case of $p_{o_2} \in B_{2,2}^*$} ]$
			\FOR{each case $S$}
				\STATE $p_{o_3} \gets$ \Call{NthOutlier}{$n_2 + 1$, $p_{o_2}$}
				\IF{$p_{o_3}$ arrives} 
					\STATE $CS\gets CS\setminus S \cup [ \text{the sub-case of $p_{o_3} \in B_{2, 1}^*$} ] \cup [ \text{the sub-case of $p_{o_3} \in B_{2,2}^*$}]$
					\FOR {each sub-case $S'$} 
						\STATE $p_{o_4} \gets$ \Call{NthOutlier}{$n_3 + 1$, $p_{o_3}$}
						\IF {$p_{o_4}$ arrives} 
							\STATE Replace $S'$ with a new candidate solution.
							\STATE $p_{o_5} \gets$ \Call{NthOutlier}{$n_4 + 1$, $p_{o_4}$}
							\IF{$p_{o_5}$ arrives}
								\STATE Abandon the solution.
							\ENDIF
						\ENDIF
					\ENDFOR
				\ENDIF
			\ENDFOR

		\ENDIF
		\STATE $\mathop{Sol}_{n_1,n_2,n_3,n_4} \gets$ the solution with the smallest larger radius, among all the sub-cases.
	\ENDFOR
	\RETURN the solution $\mathop{Sol}_{n_1,n_2,n_3,n_4}$ with the smallest radius, for all valid values of $(n_1,n_2,n_3,n_4)$.
%\EndFunction
\end{algorithmic}
\end{algorithm}

Here, \textproc{NthOutlier}$(n,p)$ returns the $n$\nth point arriving after $p$ in $P$ that does not lie in current candidate solution.

\begin{theorem}
Given $\delta^* < \alpha r^*$ and $1.2r^* \le r' < (1.2 + 2\eps/3)r^*$, Algorithm~\ref{alg:smallc} uses $\cO (dz^3)$ space and returns a $(1.8+\eps)$-approximate solution. Algorithm~\ref{alg:smallc} spends $\cO (dz^3)$ update time for each point in $P$ and answers a query in $\cO (dz^3)$.
\end{theorem}

\begin{proof}
Since our algorithm considers all possible input cases of streaming points, there is at least one feasible solution. Since every feasible solution has its larger radius at most $3r'/2$, the final solution has larger radius at most $3r'/2 \le (1.8 + \eps)r^*$.
For space complexity, our algorithm maintains at most two balls in each case, and therefore it uses $\cO(dz^3)$ space. Whenever the next point is inserted, the algorithm updates the solution for each sub-case in $\cO (d)$ time. Therefore, the algorithm spends $\cO (dz^3)$ update time for each point of $P$. Answering a query consists of choosing the minimum radius among all the candidate solutions, which amounts to $\cO (dz^3)$ time.
\end{proof}

@@

\begin{lemma}
\label{lem:2lt1}
Let $r_1^*, r_2^*$ be the radius of $1$-center and $2$-center optimal solutions with $z$ outliers, respectively. Then $r_1^* < \left(\frac{\alpha}{2} + 2\right) r_2^*$.
\end{lemma}
\begin{proof}
Suppose that $(B_{2, 1}^*(c_1^*, r_2^*), B_{2,2}^*(c_2^*, r_2^*))$ is an optimal solution for $2$-center problem. Consider $m$ as the midpoint of segment 
$c_1^*c_2^*$ (Figure~\ref{fig:2lt1}).
Clearly, $B\left(m, \frac{\delta}{2} + 2r_2^*\right)$ covers both $B_{2, 1}^*$ and $B_{2,2}^*$. So it is a solution for $1$-center problem and the proof is complete.
\end{proof}

@@

Algorithm~\ref{alg:1center} finds a $2$-approximation for $r_1^*$. By Lemma \ref{lem:1lt2} and \ref{lem:2lt1}, it gives a $2\left(\frac{\alpha}{2} + 2\right)$-approximation for $r_2^*$.

@@

Our algorithm maintains $m=\ceil{{1.2 \cdot (3 \alpha+12)}/{\eps}}$ candidate lengths.
For each candidate length $t$, assume that $r' = t$ and use it to run Algorithm~\ref{alg:smallc}. Algorithm~\ref{alg:1center} is used to estimate $r'$. Let $k_i$ be the $i$\nth non-zero radius calculated by the algorithm. Obviously, the sequence of answers given by the algorithm is increasing.

Let $l_1=k_1$ and $l_i=2^j l_1$, where $l_i$ is the smallest number satisfying $l_i \ge k_i$. By Observation~\ref{obs:2apr}, $l_i$ is a $(2\alpha + 8)$-approximation for $r_2^*$. If the interval $(0, 1.2 l_i]$ divides into $m$ equal segments (Figure~\ref{fig:findr}), then length of each segment is at most $2/3 \eps r_2^*$ and the endpoints as candidates are $\mathcal{L}_i = \set{j \times t_i \provided j = 1, \dots, m}$ where $t_i = \frac{1.2 l_i}{m}$. 
The inequality $1.2 r_2^* \le j \times l_i \le (1.2 + \frac{2\eps}{3})r_2^*$ holds for at least one of these candidates.

@@

\subsection{The Case $\delta^* \ge \alpha r^*$}
\label{subsec:bigger}

\begin{obs}
\label{obs:c+4}
Let $d$ be the distance between arbitrary pair of points from two $\alpha$-separated balls. Then $1 \le \frac{d}{\delta} \le \frac{\alpha+4}{\alpha}$.
\end{obs}

@@

\begin{cor}
\label{cor:omitting-centerpoint}
Given a set $P$ of $k(d+1)$ points in $d$\nobreakdash-dimensional Euclidean space, there is a point $c_p$, not necessarily in $P$, that any convex object that does not cover $c_p$, leaves at least $k$ points of $P$ uncovered.
\end{cor}
